---
title: " Data wrangling wrap up"
author: "Sukey"
date: "October 31, 2018"
output:
  github_document:
    toc: yes
---
# Character data

```{r, preparing work}

suppressPackageStartupMessages(library( tidyr ))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library( dplyr ))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(gtable))

library(repurrrsive)
library(robustbase)
library(gapminder)
library(stringr)



myt <- ttheme_default( 
         # Use hjust and x to center the text
         # Alternate the row fill colours
                 core = list( fg_params=list( hjust = 0.5 , x = 0.5 ),   # 1: right 0:left 0.5:center
                             bg_params=list( fill=c( "green", "pink" ) ) ),

         # Change column header to white text and red background
                 colhead = list( fg_params=list( col="white" ),
                                bg_params=list( fill="red" ) ) )
```

## 14.2.5 Exercises

* In code that doesn’t use stringr, you’ll often see paste() and paste0(). What’s the difference between the two functions? What stringr function are they equivalent to? How do the functions differ in their handling of NA?

  + The difference between `paste()` and `paste0()` is whether to use `sep`
    * paste0(..., collapse) is equivalent to paste(..., sep = "", collapse), slightly more efficiently.
    * The `usage` of this two function are as following:
```{r, results='hold'}
# paste (..., sep = " ", collapse = NULL)
# paste0(..., collapse = NULL)
paste(1:10 , sep = ",")
paste0(1:10)
```
  + They are equal to `str_c`
```{r}
str_c(1:10, sep = ",")
```
  + `paste()`  and `paste0()` will return NA as a character while ``str_c` will only return NA as it cannot connect NA value.
```{r}
paste( 1:12, c("st", "nd", NA, rep("th", 9)) ) 
paste0( 1:12, c("st", "nd", NA, rep("th", 9)) )
str_c( 1:12, c("st", "nd", NA, rep("th", 9)) )
```
  
* In your own words, describe the difference between the sep and collapse arguments to str_c().
  sep is to insert string between the input vectors while collaspe is to convert the input vectors into a single string
```{r}
str_c(letters, collapse = ",")
str_c(letters, LETTERS, collapse = ",")
str_c(letters, sep = ",")
str_c(letters, LETTERS, sep = ",")

```

* Use str_length() and str_sub() to extract the middle character from a string. What will you do if the string has an even number of characters?
```{r}
even_string <- "STAT"
odd_string <- "547"

return_middle <- function( string)
{
  if ( str_length(string) %% 2 == 0)
    index = str_length( string )/2
  else
    index = ceiling(str_length( string )/2)
  
  return( index )
}

index_even <- return_middle( even_string )
index_odd <- return_middle( odd_string )
str_sub( even_string, index_even, index_even )
str_sub( odd_string, index_odd, index_odd )

```

* What does str_wrap() do? When might you want to use it?
  + `str_wrap()`: Controls the string output format. It can wrap a long paragraph into multiple lines.
```{r}
thanks_path <- file.path(R.home("doc"), "THANKS")
thanks <- str_c(readLines(thanks_path), collapse = "\n")
( thanks <- word(thanks, 1, 3, fixed("\n\n")) )

cat(str_wrap(thanks), "\n")
```
  
* What does str_trim() do? What’s the opposite of str_trim()?
  + `str_trim` removes whitespace from start and end of string
```{r}
str.with.sapce <- "  String with trailing and leading white space\t"
(str.without.space <- str_trim( str.with.sapce ))
```
  + The oppsite of it is `str_pad`, width is the minimum width of padded strings. 
```{r}
str_pad("String with trailing and leading white space",
        side = "left",
        width = str_length(str.without.space) + 3,
        pad = " ")
```
  
* Write a function that turns (e.g.) a vector c("a", "b", "c") into the string a, b, and c. Think carefully about what it should do if given a vector of length 0, 1, or 2.
```{r}
vrctor_to_string <- function(vector)
{
  if( length(vector) == 0 )  ## if the length of vector is 0 
    return("This is an empty vector")
  
  else if( length(vector) == 1 )  ## if the length of vector is 1
    return( vector[1] )
  else
  {
    last_element <- tail( vector, n = 1 )
    other_element <- setdiff( vector, last_element )
    part1 <- str_c( other_element, collapse = "," )

    return(paste( part1, last_element, sep = " and ") )
    
  }
  
}



test1 <- c()
test2 <- c( "a" )
test3 <- c( "a", "b" )
test4 <- c( "a", "b", "c")

vrctor_to_string( test1 )
vrctor_to_string( test2 )
vrctor_to_string( test3 )
vrctor_to_string( test4 )
```

## 14.3.1.1 Exercises  

* Explain why each of these strings don’t match a \: "\", "\\", "\\\".

`\` is used as an escape character in regular expressions, how do you match a literal `\`? Well you need to escape it, creating the regular expression \\. To create that regular expression, you need to use a string, which also needs to escape `\`. That means to match a literal `\` you need to write `"\\\\"` — you need four backslashes to match one!

* How would you match the sequence `"'\`?

`"`,`'` and `\` need to be escaped and therefore, we should use`\"\'\\` to match it.
```{r}
test_string <- "\"\'\\"
writeLines ( test_string )
```

* What patterns will the regular expression \..\..\.. match? How would you represent it as a string?
```{r}
test_string1 <- "\\..\\..\\.."
writeLines( test_string1 )

```

## 14.3.2.1 Exercises  
* How would you match the literal string "$^$"?
```{r}
test_string2 <- '$^$'
writeLines( test_string2 )
```

* Given the corpus of common words in stringr::words, create regular expressions that find all words that:
  + Start with “y”.
```{r}
# str_view( stringr::words, "^y", match = TRUE)

```
  
  + End with “x”
```{r}
# str_view( stringr::words, "x$", match = TRUE )
```
  
  + Are exactly three letters long. (Don’t cheat by using str_length()!)
```{r}
# str_view( stringr::words, "^...$", match = TRUE )
```
  
  + Have seven letters or more.
```{r}
# str_view( stringr::words, ".......+", match = TRUE )
```
  
Since this list is long, you might want to use the match argument to str_view() to show only the matching or non-matching words.

## 14.3.3.1 Exercises  
* Create regular expressions to find all words that:
  * Start with a vowel.
```{r}
# str_view( stringr::words, "^[aeiou]", match = TRUE )
```
  
  * That only contain consonants. (Hint: thinking about matching “not”-vowels.)
```{r}
# str_view( stringr::words, "[aeiou]", match = FALSE)

```
  
  * End with ed, but not with eed.
```{r}
# str_view( stringr::words, "[^e]ed$", match = TRUE )
```
  
  * End with ing or ise.
```{r}
# str_view( stringr::words, "(ing)|(ise)$", match = TRUE )
```
  
* Empirically verify the rule “i before e except after c”.
```{r}
# str_view( stringr::words, "[^c]ie", match = TRUE )
```

* Is “q” always followed by a “u”?
```{r}
# str_view(stringr::words, "q[^u]", match = TRUE)
# str_view(stringr::words, "qu", match = TRUE)
```


* Write a regular expression that matches a word if it’s probably written in British English, not American English.
```{r}
x <- c("centre", "center", "metre","meter")
# str_view(x, "re$", match = TRUE)
```

* Create a regular expression that will match telephone numbers as commonly written in your country.
In my country, telephone number is usually begins with 186, 189, 158,136...
```{r}
telephone_number <- c("123456****", "186402****", "158408****", "1794***")
# str_view(telephone_number, "^(186)|(189)|(158)|(136)", match = TRUE)
```

## 14.3.4.1 Exercises
* Describe the equivalents of ?, +, * in {m,n} form.
  + ?: 0 or 1 {0,1}
  + +: 1 or more {1,}
  + *: 0 or more {0,}
* Describe in words what these regular expressions match: (read carefully to see if I’m using a regular expression or a string that defines a regular expression.)
  + ^.*$ : any string
  + "\\{.+\\}": at least one character that is in {}, e.g.{ab}
  + \d{4}-\d{2}-\d{2}: 4 digits-two digits-two digits
  + "\\\\{4}" 4`\`
* Create regular expressions to find all words that:
  + Start with three consonants.
```{r}
# str_view(stringr::words, "^[^aeiou][^aeiou][^aeiou]", match = TRUE)
```
  
  + Have three or more vowels in a row.
```{r}
# str_view(stringr::words, "[aeiou]{3,}", match = TRUE)
```
  
  + Have two or more vowel-consonant pairs in a row.
```{r}
# str_view(stringr::words, "([aeiou][^aeiou]){2}+", match = TRUE)
```
  
* Solve the beginner regexp crosswords at https://regexcrossword.com/challenges/beginner.
!()[https://github.com/STAT545-UBC-students/hw06-Sukeysun/blob/master/Data_wrangling_wrap_up_files/puzzles.jpg
]

## 14.3.5.1 Exercises
* Describe, in words, what these expressions will match:
  + (.)\1\1 : anything repeated 3 times e.g. aaa
  + "(.)(.)\\2\\1" : "item1 item2 item2 item1" e.g. p`eppe`r
  + (..)\1: "item1 item2 item1 item2" e.g.abab
  + "(.).\\1.\\1" "item1 any item item1 any item item1" e.g. b`anana`
  + "(.)(.)(.).*\\3\\2\\1" "item1 item2 item3 any item (0 or more) item item3 item2 item2" e.g. abcdcba
Construct regular expressions to match words that:
  + Start and end with the same character.
```{r}
# str_view(stringr::words, "^(.).*\\1$", match = TRUE)
```
  
  + Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)
```{r}
# str_view(stringr::words, "(.)(.).*\\1\\2", match = TRUE)
```
  
  + Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)
```{r}
# str_view(stringr::words, "(.).*\\1.*\\1", match = TRUE)
```

## 14.4.2 Exercises  

* For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.
  * Find all words that start or end with x.
```{r}
# 1
# str_view(stringr::words, "(^x)|(x$)", match = TRUE)
# 2
stringr::words[str_detect(stringr::words, "(^x)|(x$)") == TRUE]
```
  
  * Find all words that start with a vowel and end with a consonant.
```{r}
# 1
# str_view(stringr::words, "(^[aeiou])|([^aeiou]$)", match = TRUE)
# 2
head( stringr::words[str_detect(stringr::words, "(^[aeiou])|([^aeiou]$)") == TRUE] )
```
  
  * Are there any words that contain at least one of each different vowel?
```{r}
hasA = str_detect(sentences, "a")
hasE = str_detect(sentences, "e")
hasO = str_detect(sentences, "o")
hasU = str_detect(sentences, "u")
hasI = str_detect(sentences, "i")
hasY = str_detect(sentences, "y")
output.with.na <- fruit[hasA & hasE & hasO & hasU & hasI & hasY]
(output.without.na <- output.with.na[!is.na(output.with.na)])
```
  
  
* What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)
```{r, vowels,fig.height=4,fig.width=10}
counts = str_count(words, "[aeiou]")
lengths = str_length(words)
result <- tibble(words = words, counts = counts, lengths = lengths)

result1 <- result %>%
  arrange(desc(counts))

result2 <- result %>%
  mutate(prop = counts / lengths) %>%
  arrange(desc(prop))

grid.arrange(  top=textGrob( "vowels",
                            gp=gpar( fontsize = 22,font = 8 ) ),
             tableGrob( head( result1, 10 ),
             theme=myt,
             rows=NULL ),
             tableGrob( head( result2, 10 ),
             theme=myt,
             rows=NULL ),
             nrow = 1)
```

## 14.4.3.1 Exercises  

* In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.
```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colours_mod <- c("[^a-zA-Z0-9]red[^a-zA-Z0-9]", 
                 "[^a-zA-Z0-9]orange[^a-zA-Z0-9]", 
                 "[^a-zA-Z0-9]yellow[^a-zA-Z0-9]",
                 "[^a-zA-Z0-9]green[^a-zA-Z0-9]", 
                 "[^a-zA-Z0-9]blue[^a-zA-Z0-9]", 
                 "[^a-zA-Z0-9]purple[^a-zA-Z0-9]")

colour_match_mod <- str_c(colours_mod, collapse = "|")
colour_match <- str_c(colours, collapse = "|")

has_colour <- str_subset(sentences, colour_match)
matches <- str_extract_all(has_colour, colour_match)

has_colour_mod <- str_subset(sentences, colour_match_mod)
matches_mod <- str_extract_all(has_colour, colour_match_mod)

has_colour[!(has_colour %in% has_colour_mod)]
has_colour[has_colour %in% has_colour_mod]
```

* From the Harvard sentences data, extract:
  * The first word from each sentence.
```{r}
( frist.word <- str_extract( sentences, "[^ ]+" ) %>% 
  head(10) )
```
  
  * All words ending in ing.
```{r}
has_ing <- str_subset(sentences, "[^ ]+ing[ .]")
str_extract(has_ing, "[^ ]+ing[ .]") %>% str_sub(1, -2) %>% head(10)
```
  
  * All plurals.
```{r}
is_noun <- str_subset(sentences, "(a|the) [^ \\.]+")
nouns <- str_extract(is_noun, "(a|the) [^ \\.]+")
is_plural <- str_subset(nouns, "(a|the) [^ \\.\\']+(s|\\')$")
str_extract(is_plural, "(a|the) [^ \\.\\']+(s|\\')$")

```
  
## 14.4.4.1 Exercises

* Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.
* Find all contractions. Separate out the pieces before and after the apostrophe.
```{r}
## 1
numbers = c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten")
numbers_match = str_c("(", str_c(numbers, collapse = "|"), ") [^ .]+")

has_number <- str_subset(sentences, numbers_match)
str_extract(has_number, numbers_match) %>% head(10)

## 2
has_contraction = str_subset(sentences, "[^ ]+\\'[^ .]+")
str_extract(has_contraction, "[^ ]+\\'[^ .]+")
```

## 14.4.5.1 Exercises  
* Replace all forward slashes in a string with backslashes.
```{r}
str = "STAT/547/HW06"
writeLines(str)
writeLines(str_replace_all(str, "/", "\\\\"))
```

* Implement a simple version of str_to_lower() using replace_all().
```{r}
str_replace_all(LETTERS, c("A" = "a", "B" = "b", "C" = "c"))
```

* Switch the first and last letters in words. Which of those strings are still words?
```{r}
switched <- str_replace(words, "^(.)(.*)(.)$", "\\3\\2\\1")
words[words %in% switched] %>%  head()

```

## 14.4.6.1 Exercises  
* Split up a string like "apples, pears, and bananas" into individual components.
```{r}
str_split("apples, pears, and bananas", ", and |, ")
```

* Why is it better to split up by boundary("word") than " "?
boundary(word) deals with whitespaces (any and any number)

* What does splitting with an empty string ("") do? Experiment, and then read the documentation.
Split by character (equivalent to boundary("character"))

## 14.5.1 Exercises
How would you find all strings containing \ with regex() vs. with fixed()?
```{r}
strings = c("STAT\\", "547\\", "HW06")
# str_view(strings, regex("\\\\"))
# str_view(strings, fixed("\\"))
```

What are the five most common words in sentences?
```{r, five most common words in sentences, fig.height=4,fig.width=10}
words <- str_split(sentences, boundary("word"), simplify = T) %>% 
  as.vector()
  
result <- tibble(words = str_to_lower(words)) %>%
  group_by(words) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  head(6) 

grid.arrange(  top=textGrob( "five most common words in sentences",
                            gp=gpar( fontsize = 22,font = 8 ) ),
             tableGrob( result,
             theme=myt,
             rows=NULL ) )

```

## 14.7.1 Exercises
* Find the stringi functions that:
  * Count the number of words. `stri_count_words`
  * Find duplicated strings. `stri_duplicated`
  * Generate random text. `stri_rand_strings (stri_rand_lipsum for Lorem ipsum)`
How do you control the language that stri_sort() uses for sorting?
By the locale argument.

# Writing functions
Write one (or more) functions that do something useful to pieces of the Gapminder or Singer data. It is logical to think about computing on the mini-data frames corresponding to the data for each specific country, location, year, band, album, … This would pair well with the prompt below about working with a nested data frame, as you could apply your function there. 




I would like to create a function to calculate the sum of the gdp of a specific continent in all years. The input of the funcion is:
  * Continent: the specific continent which I want to compute, the default continent is Oceania
  
```{r, compute the gdp of a given continent}
explore_gdp <- 
  function( Continent = "Oceania")
  {
    gapminder %>% 
      ## choose the continent and the year you are interested in
      filter (continent == Continent) %>% 
      group_by( year ) %>% 
      ## compute the gdp of each country in the given year of the input continent
      mutate ( gdp = gdpPercap * pop ) %>% 
      summarize( sumOfGdp = sum( gdp ) ) %>% 
      mutate ( continent = Continent)
      # str_c("The total gdp of", Continent, "in", Year, "is", totalgdp, "dollars", sep = " " )
  }
```

This function should return the sum of gdp of a specific continent in a given year. Let's check if it works.
```{r, total Gdp of Oceania, fig.height=4,fig.width=10}
gdp_Oceania <- explore_gdp( Continent = "Oceania")

grid.arrange(  top=textGrob( "Total Gdp of Oceania",
                            gp=gpar( fontsize = 22,font = 8 ) ),
             tableGrob( head( gdp_Oceania, 10 ),
             theme=myt,
             rows=NULL ) )
```
As shown in the table above, the total gdp of Oceania in 1957 is 133653656026.872 dollars

The `explore_gdp` function could only give me the total gdp of a given continent in a given year that are shown in the gapminder dataset. However, I also want to know the predicted data of all continents during some years that are not in gapminder dataset.
First, let's check the relationship between year and total gdp in all continents
```{r, check the relationship between year and total gdp in all continents,fig.height=6,fig.width=12}
gdp_gapminder <- gapminder %>% 
  ## add a new column called totalgdp
  mutate( totalgdp = gdpPercap*pop )
  
gdp_gapminder %>% 
  ggplot()+
  facet_wrap(~continent)+
  geom_point( aes( year,log( totalgdp), colour = continent ) ) +
  theme( strip.text.x = element_text( size = 10 , angle = 30),
          strip.text.y = element_text( size = 10, face = "bold" ),
          strip.background = element_rect( colour = "red", fill = "#CCCCFF" )) +
  labs( title = "  total Gdp for each continent  ") +
 theme_bw()+
   theme(
         axis.text.x = element_text( size = 10 ),
         axis.text.y = element_text( size = 10 ),
         axis.title = element_text( size = 14),
         plot.background = element_rect( fill = "white", colour = "black" ),
         plot.title = element_text( size = 15,hjust = 0.5),
         panel.border = element_rect(linetype = "dashed", fill = NA)
         )


  

```
As we can see in the graph above, the log of total gdp increases with year and there seems to be a linear regression between them.
Now let's make a prediction function

```{r,Predicted total Gdp of given years,fig.height=5,fig.width=12}
predict_gdp <- 
  function( Year = c(1952,1953,1954, 1955, 1956) )
  {
    newyear <- tibble(year = Year)
    ## call the explore_gdp I created before to get the total gdp of a given continent
    continents = unique(gapminder$continent)
    
    #gdp <- data.frame()
    predict_data <- data.frame()
    for (con in continents)
    {
      
      temp_data <- explore_gdp( Continent = con)
      
      ## use robust regression to train data
      temp_lm <- lmrob( log(sumOfGdp) ~ year, data = temp_data)
      
      ## create a predicted dataframe
      temp_predict <- tibble(
        continent = con,
        year = Year,
        predict.log.gdp = predict( temp_lm, newdata = newyear)
        
      )
      
      ## paste the predicted data row by row
      predict_data <- bind_rows(predict_data, temp_predict)
    }
    
    return (predict_data)
  }



newyear_Oceania <- predict_gdp( Year = c(1950,1960,1970,1980))

plt_Oceania <- newyear_Oceania %>% 
  ggplot()+
  
  ## plot the data separately according to continent.
  facet_wrap(~continent)+
  geom_point( aes( year,predict.log.gdp, colour = continent ) ) +
  theme( strip.text.x = element_text( size = 10 , angle = 30),
          strip.text.y = element_text( size = 10, face = "bold" ),
          strip.background = element_rect( colour = "red", fill = "#CCCCFF" )) +
  labs( title = "  total Gdp for each continent  ") +
 theme_bw()+
   theme(
         axis.text.x = element_text( size = 10 ),
         axis.text.y = element_text( size = 10 ),
         axis.title = element_text( size = 10),
         plot.background = element_rect( fill = "white", colour = "black" ),
         plot.title = element_text( size = 15,hjust = 0.5),
         panel.border = element_rect(linetype = "dashed", fill = NA)
         )


grid.arrange( top =textGrob( " Predicted total Gdp of given years ",
                            gp=gpar( fontsize = 20,font = 8 ) ),
             tableGrob( head( newyear_Oceania,10 ),
             theme=myt,
             rows=NULL ),
             plt_Oceania,
             nrow =1)




```




